The Effects of Varying Depth in Cosmic Shear Surveys

### 2019-05-07 model_xipm.tex
First rework, now including the new insights. Probably too long.

\section{Modelling the shear correlation functions}
\label{sec:xipm}
A convenient way to infer cosmological information from observational data are the shear correlation functions $\xi_\pm$, which are defined as \[
\xi_\pm = \la \gamma_{\rm t}\gamma_{\rm t}\ra \pm \la \gamma_\times\gamma_\times\ra \, .
\]
They are the prime estimators to quantify a cosmic-shear signal since it is simple to include a weighting of the shear-measurements into the correlation functions and, contrary to the power spectrum, one does not have to worry about the shape of the survey footprint, or masked regions. Cosmologically, given two comoving distance probability distributions of sources $p_i(\chi),p_j(\chi)$, one can compute the shear correlation function from the underlyting matter power spectrum $P_\delta$ via \todo{Citation!}\begin{align}
\label{eq:xipm-pkappa}
\xi_\pm[\theta,p_i,p_j] =& \int_0^\infty \frac{{\rm d}l\,l}{2\pi}J_{0,4}(l\theta)P(l,p_i,p_j)\, , \\
\label{eq:pkappa-pdelta/lenseff}
P(l,p_i,p_j) =& \frac{9 H_0^4\Omega_{\rm m}^2}{4c^4}\int_0^{\chi_h} {\rm d}\chi\frac{g(\chi,p_i)g(\chi,p_j)}{a^2(\chi)}P_\delta\left(\frac{l}{f_K(\chi)},\chi\right)\, , \\
\label{eq:lenseff}
g(\chi,p_i) =& \int_\chi^{\chi_H} {\rm d}\chi'\, p_i(\chi') \frac{f_K(\chi'-\chi)}{f_K(\chi')}\, .
\end{align}
Here, $J_{0,4}$ denote the 0-th and 4-th order Bessel Functions.
\subsection{Using an analytic Model}
\label{sec:xipm_analytic}
For a first simple analysis we will assume that a deeper redshift distribution just yields a stronger shear signal, in the sense that the shear field for a deeper redshift distribution gets multiplied by a weight $W$. While this is not true for a 3-dimensional matter distribution, it should be valid for small variations in redshift. Additionally, we assume that a higher depth does not only lead to a stronger average shear, but also to a higher galaxy number density, implying a correlation between those two quantities.

Let $N^i(\b \theta),N^j(\b\theta)$ be the average weighted number of galaxies per pointing in redshift bins $i$ and $j$ and let $W^i(\b \theta),W^j(\b\theta)$ be the weighting of average shear. The observed correlation function $\xi^{ij,\text{obs}}_\pm(\theta)$ now changes from one of constant depth $\xi_\pm^{ij,\rm{const}}(\theta)$ via (compare Eq. \todo{Take eq. from Hildebrandt et al. (2017)?})
\begin{align}
\xi^{ij,\text{obs}}_\pm(\theta) = & \frac{\la N^i(\bth)N^j(\bthp)\gamma^{i,\rm{obs}}_{\rm t}(\bth)\gamma^{j,\rm{obs}}_{\rm t}(\bthp)\ra }{\la N^i(\bth)N^j(\bthp)\ra} \pm \frac{\la N^i(\bth)N^j(\bthp)\gamma^{i,\rm{obs}}_\times(\bth)\gamma^{j,\rm{obs}}_\times(\bthp)\ra }{\la N^i(\bth)N^j(\bthp)\ra} \nonumber\\
 = & \frac{\la N^i(\bth)N^j(\bthp)W^i(\bth)W^j(\bthp)\ra}{\la N^i(\bth)N^j(\bthp)\ra} \xi_{\pm}^{ij,\rm{const}}(\theta) \, ,
 \label{eq:xipmblub1}
 \end{align}
 where the average $\la\ldots\ra$ represents both an ensemble average as well as an average over the position $\bth$.
 Assuming that depth and galaxy number density of neighbouring pointings are uncorrelated, the only important property of a galaxy pair is whether or not they lie in the same pointing, which is described by the function $E(\theta)$.

@@

Ignoring correlations higher than second order\footnote{It is not inherently obvious that this is a valid assumption. However, after performing both calculations we noticed no difference between the outcomes of both equations}, and performing the same calculation for the denominator of Eq.\,\eqref{eq:xipmblub1}, we get

@@

The analysis of data from the Kilo-Degree Survey showed that the redshift-distribution of sources was highly correlated with the depth in the $r$-band. We thus chose to separate the survey into 10 percentiles, sorted by $r$-band depth, meaning that if a pointing had a worse depth than 90\% of the other pointings, it would belong to the first percentile, and so on. Now for each percentile $m$ and each tomographic redshift bin $i$ we can extract a weighted number of galaxies $N^i_m$ and, in case the pointing overlaps with a spectroscopic survey, a source redshift distribution $p^i_m(z)$. Using \eqref{eq:xipm-pkappa}, we can compute the shear correlation functions $\xi_{\pm,mn}^{ij}(\theta)$ for each set of percentiles $m,n$ and redshift bins $i,j$\footnote{For the calculation of the shear correlation functions we use the \textsc{Nicaea}-program. Among other things, it calculates the shear correlation functions for a given cosmology and source redshift distribution. To estimate the power spectrum on nonlinear scales we use the methods developed by \citet{2012ApJ...761..152T}.}. When we compute the measured shear correlation functions of a survey, we take the weighted average of tangential and cross shears of all pairs of galaxies (\citet{2017MNRAS.465.1454H} give a good overview for the process\todo{Cite the same eq. as above from Hildebrandt et al. (2017)}). If, for a single pair of galaxies, one galaxy lies in the $m$-th percentile of redshift bin $i$ and the second one lies in the $n$-th percentile of redshift bin $j$, then their contribution to the observed correlation functions is, on average, $\xi_{\pm,mn}^{ij}(\theta)$. This means that if we know each of those single correlation functions, we can reconstruct the total correlation functions via a weighted average of the single functions. Formally, we define \[
\xi_\pm^{ij,\rm{obs}}(\theta) = \frac{\sum_{m,n} P_{mn}^{ij}(\theta)\xi_{\pm,mn}^{ij}(\theta)}{\sum_{m,n} P_{mn}^{ij}(\theta)}\, ,

@@

In this analysis, we will assume an infinitely large survey footprint with an uncorrelated distribution of depth. We will later discuss the validity of these assumptions as well as possible mitigation strategies. For $m\neq n$ we know that the pair of galaxies has to lie in different pointings, which is accounted for by including the factor $[1-E(\theta)]$. Furthermore, the first galaxy has to lie in percentile $m$, the probability of which is $1/10$. When the first pointing is of percentile $m$, the probability that a galaxy in a different pointing is of percentile $n$ is also equal to $1/10$. The impact of such a galaxy pair on the correlation functions scales with the product of the weighted number of galaxies $N_m^i,N_n^j$. We get for $n\neq m$: \[
P_{mn}^{ij}(\theta) = [1-E(\theta)]\frac{1}{100} N_m^i N_n^j\, .
\label{eq:pmnij_corr1}
\]
For the calculation of $P_{mm}^{ij}(\theta)$ we have to account for a different possibility: In case that the galaxy lies in the same pointing [accounted for by the factor $E(\theta)$], it automatically is of the same percentile. We therefore set \[
P_{mm}^{ij}(\theta) = E(\theta)\frac{1}{10} N_m^iN_m^j + [1-E(\theta)]\frac{1}{100} N_m^i N_m^j \, .
\label{eq:pmnij_corr2}
\]
We can then write $P_{mn}^{ij}(\theta)$ as: \[
P_{mn}^{ij}(\theta) = E(\theta)\frac{1}{10} N_m^iN_n^j\,\delta_{mn} + [1-E(\theta)]\frac{1}{100} N_m^i N_m^j \, ,
\label{eq:pmnij_uncorr}
\]
where $\delta_{mn}$ denotes the Kronecker delta.
Inserting this into Eq.\,\eqref{eq:def_xiobs}, we compute
\begin{align}
\xi_{\pm,mn}^{ij,\rm{obs}}(\theta) = & \frac{1}{C}\sum_{m=1}^{10} N_m^i \left\{ E(\theta) N_m^j \xi_{\pm,mm}^{ij}(\theta) + \frac{\big[1-E(\theta)\big]}{10}\sum_{n=1}^{10}N_n^j \xi_{\pm,mn}^{ij}(\theta)\right\}\, ,
\label{eq:correctionfunction1}
\end{align}
with the normalization
\[
C = \sum_{m=1}^{10} N_m^i \bigg[ E(\theta)  N_m^j + \frac{\big[1-E(\theta)\big]}{10}\sum_{n=1}^{10} N_n^j\bigg]\, .
\]
A more mathematically rigorous derivation of this function can be found in Appendix \ref{sec:calc of xipm}.

If we want to compute this for all 5 redshift bins of the KV450-survey, this forces us to calculate 1275 correlation functions and add them, thus yielding potential numerical errors (apart from being computationally expensive). However, if we examine Eq.\,\eqref{eq:lenseff}, we see that the comoving distance distribution of sources factors in linearly. This in turn implies that in Equations \eqref{eq:pkappa-pdelta/lenseff} and \eqref{eq:xipm-pkappa} both source distance distributions factor in linearly. This basically means that, instead of adding correlation functions, we can add their respective redshift distributions and compute the correlation functions of that. In particular, we can define the \textit{combined number of galaxies} $N^i$ and \textit{average redshift distribution} $p^i(z)$ of tomographic bin $i$ as \[
N^i\equiv\sum_m N_m^i\, , \qquad p^i(z) = \frac{\sum_m N_m^i p_m^i(z)}{\sum_m N_m^i} \, .
\]
If we define $\xi^{ij}_\pm$ as the correlation functions between the average redshift distributions $p^i(z)$ and $p^j(z)$, then we observe: \[
\sum_{m,n}N_m^iN_n^j\xi^{ij}_{\pm,mn} = N^iN^j\xi^{ij}_\pm\, .
\]
Consequently, we can apply this to \eqref{eq:correctionfunction1}, yielding
\begin{equation}
\xi_{\pm}^{ij,\rm{obs}}(\theta) = \frac{1}{C}\left\{ E(\theta)\left[\sum_{m=1}^{10} N_m^iN_m^j \xi_{\pm,mm}^{ij}(\theta)\right] +\frac{\big[1-E(\theta)\big]}{10}\xi_\pm^{ij}(\theta)N^iN^j\right\}\, .
\label{eq:correctionfunction2}
\end{equation}
For each set of redshift bins we thus only have to compute eleven correlation functions, which reduces the number of functions to compute from 1275 to 165. We can see that for large distances $\theta$, such that $E(\theta)=0$ holds, the observed correlation functions agree with the ones that we would usually calculate.

>>>>

\section{Modelling the shear correlation functions}
\label{sec:xipm}
A convenient way to infer cosmological information from observational data are the shear correlation functions $\xi_\pm$, which are defined as \[
\xi_\pm = \la \gamma_{\rm t}\gamma_{\rm t}\ra \pm \la \gamma_\times\gamma_\times\ra \, .
\]
They are the prime estimators to quantify a cosmic-shear signal since it is simple to include a weighting of the shear-measurements into the correlation functions and, contrary to the power spectrum, one does not have to worry about the shape of the survey footprint, or masked regions. Cosmologically, given two comoving distance probability distributions of sources $p_i(\chi),p_j(\chi)$, one can compute the shear correlation function from the underlyting matter power spectrum $P_\delta$ via \todo{Citation!}\begin{align}
\label{eq:xipm-pkappa}
\xi_\pm^{ij}(\theta) =& \int_0^\infty \frac{{\rm d}\ell\,\ell}{2\pi}J_{0,4}(\ell\theta)P^{ij}(\ell)\, , \\
\label{eq:pkappa-pdelta/lenseff}
P^{ij}(\ell) =& \frac{9 H_0^4\Omega_{\rm m}^2}{4c^4}\int_0^{\chi_h} {\rm d}\chi\frac{g^i(\chi)g^j(\chi)}{a^2(\chi)}P_\delta\left(\frac{\ell}{f_K(\chi)},\chi\right)\, , \\
\label{eq:lenseff}
g^i(\chi) =& \int_\chi^{\chi_H} {\rm d}\chi'\, p_i(\chi') \frac{f_K(\chi'-\chi)}{f_K(\chi')}\, .
\end{align}
Here, $J_{0,4}$ denote the 0-th and 4-th order Bessel Functions.
\subsection{Using an analytic Model}
\label{sec:xipm_analytic}
For a first simple analysis we will assume that a deeper redshift distribution just yields a stronger shear signal, in the sense that the shear field for a deeper redshift distribution gets multiplied by a weight $W$. While this is not true for a 3-dimensional matter distribution, it should be valid for small variations in redshift. Additionally, we assume that a higher depth does not only lead to a stronger average shear, but also to a higher galaxy number density, implying a correlation between those two quantities.

Let $N^i(\b \theta),N^j(\b\theta)$ be the average weighted number of galaxies per pointing in redshift bins $i$ and $j$ and let $W^i(\b \theta),W^j(\b\theta)$ be the weighting of average shear. The observed correlation function $\xi^{ij,\text{obs}}_\pm(\theta)$ now changes from one of constant depth $\xi_\pm^{ij,\rm{const}}(\theta)$ via (compare Eq. \todo{Take eq. from Hildebrandt et al. (2017)?})
\begin{align}
\xi^{ij,\text{obs}}_\pm(\theta) = & \frac{\la N^i(\bth)N^j(\bthp)\gamma^{i,\rm{obs}}_{\rm t}(\bth)\gamma^{j,\rm{obs}}_{\rm t}(\bthp)\ra }{\la N^i(\bth)N^j(\bthp)\ra} \pm \frac{\la N^i(\bth)N^j(\bthp)\gamma^{i,\rm{obs}}_\times(\bth)\gamma^{j,\rm{obs}}_\times(\bthp)\ra }{\la N^i(\bth)N^j(\bthp)\ra} \nonumber\\
 = & \frac{\la N^i(\bth)N^j(\bthp)W^i(\bth)W^j(\bthp)\ra}{\la N^i(\bth)N^j(\bthp)\ra} \xi_{\pm}^{ij,\rm{const}}(\theta) \, ,
 \label{eq:xipmblub1}
 \end{align}
 where the average $\la\ldots\ra$ represents both an ensemble average as well as an average over the position $\bth$.
 Assuming that depth and galaxy number density of neighbouring pointings are uncorrelated, the only important property of a galaxy pair is whether or not they lie in the same pointing. We want to denote the probability, that a random galaxy pair of separation $\theta$ lies in the same pointing with $E(\theta)$. This function is depicted in Figure \ref{fig:eoftheta_lin}; an exact equation and derivation are given in \ref{sec:model_e}.

\begin{figure}
 \centering
 \includegraphics[width=0.5\textwidth]{images/eoftheta.pdf}
 \caption{Probability that a random pair of galaxies of distance $\theta$ lie in the same pointing.}
 \label{fig:eoftheta_lin}
\end{figure}

@@

Ignoring correlations higher than second order\footnote{It is not inherently obvious that this is a valid assumption. However, after performing both calculations we noticed no difference between the outcomes of both equations.}, and performing the same calculation for the denominator of Eq.\,\eqref{eq:xipmblub1}, we get

@@

The analysis of data from the Kilo-Degree Survey showed that the redshift-distribution of sources was highly correlated with the depth in the $r$-band. We thus chose to separate the survey into 10 percentiles, sorted by $r$-band depth, meaning that if a pointing had a worse depth than 90\% of the other pointings, it would belong to the first percentile, and so on. For each percentile $m$ and each tomographic redshift bin $i$ we can extract a weighted number of galaxies $N^i_m$ and, in case the pointing overlaps with a spectroscopic survey, a source redshift distribution $p^i_m(z)$. Using \eqref{eq:xipm-pkappa}, we can compute the shear correlation functions $\xi_{\pm,mn}^{ij}(\theta)$ for each set of percentiles $m,n$ and redshift bins $i,j$\footnote{For the calculation of the shear correlation functions we use the \textsc{Nicaea}-program. Among other things, it calculates the shear correlation functions for a given cosmology and source redshift distribution. To estimate the power spectrum on nonlinear scales we use the methods developed by \citet{2012ApJ...761..152T}.}. When we compute the measured shear correlation functions of a survey, we take the weighted average of tangential and cross shears of all pairs of galaxies (\citet{2017MNRAS.465.1454H} give a good overview for the process\todo{Cite the same eq. as above from Hildebrandt et al. (2017)}). If, for a single pair of galaxies, one galaxy lies in the $m$-th percentile of redshift bin $i$ and the second one lies in the $n$-th percentile of redshift bin $j$, then their contribution to the observed correlation functions is, on average, $\xi_{\pm,mn}^{ij}(\theta)$. This means that if we know each of those single correlation functions, we can reconstruct the total correlation functions via a weighted average of the single functions. Formally, we define \[
\xi_\pm^{ij,\rm{obs}}(\theta) = \frac{\sum_{m,n} P_{mn}^{ij}(\theta)\,\xi_{\pm,mn}^{ij}(\theta)}{\sum_{m,n} P_{mn}^{ij}(\theta)}\, ,

@@

In this analysis, we will assume an infinitely large survey footprint with an uncorrelated distribution of depth. We will later discuss the validity of these assumptions as well as possible mitigation strategies. For $m\neq n$ we know that the pair of galaxies has to lie in different pointings, which is accounted for by including the factor $[1-E(\theta)]$. Furthermore, the first galaxy has to lie in percentile $m$, the probability of which is $1/10$. The pointing of the second galaxy has to be of percentile $n$; the probability of that is also equal to $1/10$. The impact of such a galaxy pair on the correlation functions scales with the product of the weighted number of galaxies $N_m^i,N_n^j$. We get for $n\neq m$: \[
P_{mn}^{ij}(\theta) = [1-E(\theta)]\frac{1}{100} N_m^i N_n^j\, .
\label{eq:pmnij_corr1}
\]
For the calculation of $P_{mm}^{ij}(\theta)$ we have to account for a different possibility: In case that the galaxy lies in the same pointing [accounted for by the factor $E(\theta)$], it automatically is of the same percentile. We therefore set \[
P_{mm}^{ij}(\theta) = E(\theta)\frac{1}{10} N_m^iN_m^j + [1-E(\theta)]\frac{1}{100} N_m^i N_m^j \, .
\label{eq:pmnij_corr2}
\]
We can then write $P_{mn}^{ij}(\theta)$ as: \[
P_{mn}^{ij}(\theta) = E(\theta)\frac{1}{10} N_m^iN_n^j\,\delta_{mn} + [1-E(\theta)]\frac{1}{100} N_m^i N_n^j \, ,
\label{eq:pmnij_uncorr}
\]
where $\delta_{mn}$ denotes the Kronecker delta.
Inserting this into Eq.\,\eqref{eq:def_xiobs}, we compute
\begin{align}
\xi_{\pm,mn}^{ij,\rm{obs}}(\theta) = & \frac{1}{C}\sum_{m=1}^{10} N_m^i \left\{ E(\theta) N_m^j \xi_{\pm,mm}^{ij}(\theta) + \frac{\big[1-E(\theta)\big]}{10}\sum_{n=1}^{10}N_n^j \xi_{\pm,mn}^{ij}(\theta)\right\}\, ,
\label{eq:correctionfunction1}
\end{align}
with the normalization
\[
C = \sum_{m=1}^{10} N_m^i \bigg[ E(\theta)  N_m^j + \frac{\big[1-E(\theta)\big]}{10}\sum_{n=1}^{10} N_n^j\bigg]\, .
\]
A more mathematically rigorous derivation of this function can be found in Appendix \ref{sec:calc of xipm}.

If we want to compute this for all 5 redshift bins of the KV450-survey, this forces us to calculate 1275 correlation functions and add them, thus yielding potential numerical errors (apart from being computationally expensive). However, if we examine Eq.\,\eqref{eq:lenseff}, we see that the comoving distance distribution of sources factors in linearly. This in turn implies that in Equations \eqref{eq:pkappa-pdelta/lenseff} and \eqref{eq:xipm-pkappa} both source distance distributions factor in linearly. This basically means that, instead of adding correlation functions, we can add their respective redshift distributions and compute the correlation functions of that. In particular, we can define the \textit{combined number of galaxies} $N^i$ and \textit{average redshift distribution} $p^i(z)$ of tomographic bin $i$ as \[
N^i\equiv\sum_m N_m^i\, , \qquad p^i(z) = \frac{\sum_m N_m^i p_m^i(z)}{\sum_m N_m^i} \, .
\]
If we define $\xi^{ij}_\pm$ as the correlation functions between the average redshift distributions $p^i(z)$ and $p^j(z)$, then we observe: \[
\sum_{m,n}N_m^iN_n^j\xi^{ij}_{\pm,mn} = N^iN^j\xi^{ij}_\pm\, .
\]
Consequently, we can apply this to \eqref{eq:correctionfunction1}, yielding
\begin{equation}
\xi_{\pm}^{ij,\rm{obs}}(\theta) = \frac{1}{C}\left\{ E(\theta)\left[\sum_{m=1}^{10} N_m^iN_m^j \xi_{\pm,mm}^{ij}(\theta)\right] +\frac{\big[1-E(\theta)\big]}{10}\xi_\pm^{ij}(\theta)N^iN^j\right\}\, .
\label{eq:correctionfunction2}
\end{equation}
For each set of redshift bins we thus only have to compute eleven correlation functions, which reduces the number of functions to compute from 1275 to 165.

### 2019-05-07 paperdraft.tex
Modified derivation of correlation function
Added Appendix
First semi-complete version after recent changes. Will still be subject to many modifications.
First rework, now including the new insights. Probably too long.

\renewcommand{\rm}{\mathrm}
\def\inv{^{-1}}
\def\({\left(}
\def\r){\right)}
\def\b#1{\bm{#1}}
\def\Dd{D_{\text{d}}}
\def\Dds{D_{\text{ds}}}
\def\Ds{D_{\text{s}}}
\def\la{\left<}
\def\ra{\right>}
\def\gammao{\gamma^{\text{obs}}}
\def\gammaoh{\hat{\gamma}^{\text{obs}}}
\def\epsilonsij{\epsilon_{a_i}\epsilon_{b_j}^*}
\def\epsilonsii{\epsilon_{a_i}\epsilon_{b_i}^*}
\def\deltafuncij{\Delta(\b\theta_{a_i}-\b\theta_{b_j})}
\def\d{\rm{d}}

@@

\usepackage{bm}
\usepackage{verbatim}
%
\usepackage{graphicx}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{txfonts}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage[options]{hyperref}
% To add links in your PDF file, use the package "hyperref"
% with options according to your LaTeX or PDFLaTeX drivers.
%
\usepackage{todonotes}

\begin{document}

@@

  \abstract
% context (optional)
   {Cosmic shear proves to be a powerful tool to study the properties of the local Universe. The discrepancy in the parameter $S_8$ between measurements in the local Universe and the Cosmic Microwave Background motivates further investigation of yet unaccounted systematic biases. 
% aims
   Especially ground-based surveys are subject to a variation in depth. We want to understand and quantify the resulting effects. In particular, we check if they introduce a bias to the cosmological parameters and if they can be responsible for the occurence of $B$-Modes.
% methods
   We construct a semi-analytic model to estimate the impact on the shear correlation function and analyze the implications for cosmological parameters. Furthermore we construct COSEBIs of the correlation functions to quantify the occuring B-Modes.
% results
   For the Kilo-Degree Survey this effect introduces an error in $\xi_\pm$ of the order of a few percent on small scales, which is responsible for a $0.1\sigma$ bias in $\Omega_{\rm m}$ and $\sigma_8$. However, the parameter $S_8$ is robust against this modification. We also report the occurrence of B-Modes, although not to a significant degree.
% conclusions (optional)
   We conclude that the effects of varying optical depth for ground-based surveys on cosmological parameters are not yet significant, but should be accounted for in next-generation experiments.}

@@

\bibliographystyle{aa}
\bibliography{cite}
\begin{appendix}

\include{appendix}

\end{appendix}
\end{document}
%

>>>>

\renewcommand{\rm}{\mathrm}
\def\inv{^{-1}}
\def\({\left(}
\def\r){\right)}
\def\b#1{\bm{#1}}
\def\Dd{D_{\text{d}}}
\def\Dds{D_{\text{ds}}}
\def\Ds{D_{\text{s}}}
\def\la{\left<}
\def\ra{\right>}
\def\gammao{\gamma^{\text{obs}}}
\def\gammaoh{\hat{\gamma}^{\text{obs}}}
\def\epsilonsij{\epsilon_{a_i}\epsilon_{b_j}^*}
\def\epsilonsii{\epsilon_{a_i}\epsilon_{b_i}^*}
\def\deltafuncij{\Delta(\b\theta_{a_i}-\b\theta_{b_j})}
\def\d{\rm{d}}
\def\bthp{\b \theta' + \b \theta}
\def\bth{\b \theta'}

@@

\usepackage{bm}
\usepackage{verbatim}
%
\usepackage{graphicx}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{txfonts}
\usepackage{subcaption}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage[options]{hyperref}
% To add links in your PDF file, use the package "hyperref"
% with options according to your LaTeX or PDFLaTeX drivers.
%
\usepackage{todonotes}
%\renewcommand{\todo}{\emph}
\begin{document}

@@

  \abstract
% context (optional)
   {Cosmic shear proves to be a powerful tool to study the properties of the local Universe. The discrepancy in the parameter $S_8$ between measurements in the local Universe and the Cosmic Microwave Background motivates further investigation of yet unaccounted systematic biases. 
% aims
   Especially ground-based surveys are subject to a variation in depth. We want to understand and quantify the resulting effects. In particular, we check whether they introduce a bias to the cosmological parameters and if they can be responsible for the occurence of $B$-Modes.
% methods
   We construct a semi-analytic model to estimate the impact on the shear correlation function and analyze the implications for cosmological parameters. Furthermore we construct COSEBIs of the correlation functions to quantify the occuring B-Modes.
% results
   For the Kilo-Degree Survey this effect introduces an error in $\xi_\pm$ of the order of a few percent on small scales, which is responsible for a $0.1\sigma$ bias in $\Omega_{\rm m}$ and $\sigma_8$. However, the parameter $S_8$ is robust against this modification. We also report the occurrence of B-Modes, although not to a significant degree.
% conclusions (optional)
   We conclude that the effects of varying depth for ground-based surveys on cosmological parameters are not yet significant, but should be accounted for in next-generation experiments.}

@@

\bibliographystyle{aa}
\bibliography{cite}
\begin{appendix}
\include{appendix}
\end{appendix}
\end{document}
%

### 2019-05-07 results.tex
First rework, now including the new insights. Probably too long.

\section{Results}
\label{sec:results}
We applied both our methods to data from the KV450 survey and computed the ratio of observed and modeled correlation functions. Furthermore, we conducted numerical simulations investigating the same issue: A $100\,\rm{deg}^2$ field in the Scinet Light Cone Simulations (SLICS) \citep{2018MNRAS.481.1337H} was randomly separated into 10 percentiles. For each tomographic redshift bin of each percentile, galaxies were placed to trace the respective redshift distribution. Afterwards, their expected shear was determined (shape-noise in the form of intrinsic ellipticities of galaxies was not included). This was compared to a set of simulations where the galaxies were simply distributed according to the combined redshift distribution of each respective tomographic bin\todo{Catherine, is this what you did? If you are unhappy with any of these explanations, or feel like I missed something, please feel free to add it :)}. As in \citet{2018arXiv181206076H}, we have separated the data in 5 tomographic redshift bins and performed our analysis for a cross-correlation of all bins. The result can be seen in Figure \ref{fig:all_xis}.
	\begin{figure}
	\centering
	\includegraphics[width=1\textwidth]{images/allxis_0111.pdf}
	\caption{The ratio of observed to modeled correlation functions for the analytic method (green), the semi-analytic method (blue) and the numerical simulations (red) for a cross-correlation of all redshift bins.}
	\label{fig:all_xis}
	\end{figure}
We can see that for high redshift bins, the analytic and the semi-analytic methods are consistent, whereas for low redshift bins they significantly diverge. We explain this due to the facts that the analyic method uses simplifications that are redshift-dependent and only hold for small variations in redshift, which is not fulfilled in the low redshift bins.\todo{Should I include the plot number density vs average redshift from the talk? Maybe in the appendix to avoid too many Figures?}\footnote{We also observe that in the semi-analytic model, $\xi_-$ seems to be much stronger affected by this effect: Following Equation \eqref{eq:xipm-pkappa}, $\xi_+$ is computed by filtering the power spectrum with the 0-th order Bessel function. This function peaks at $\ell\theta=0$, meaning that for all values of $\theta$, the correlation function $\xi_+$ is sensitive to small values of $\ell$, corresponding to large separations $\theta$. However, $\xi_-$ is obtained by filtering with the 4-th order Bessel function, which peaks at approximately $\ell\theta\approx 5$, so for different $\theta$ this function is sensitive to varying parts of the convergence power spectrum. A more detailed analysis of this can be found in the Appendix of \citet{2017MNRAS.471.4412K}.}

The simulations seem to be in rough agreement with the models, but there are some features that can not be explained. Especially we observe that the ratios of the correlation functions for large values of $\theta$ seem to consistently surpass unity, which can not be explained by our models. The simulations seem to be in rough agreement with the models, but there are some significant differences. After a thorough analysis we explain these discrepancies the following way: The simulations were performed on a $100\,\rm{deg}^2$ field, which means that shot-noise of the fields plays a significant role. After performing the same simulations for a different distribution of depth between the pointings and obtaining completely different results, we are quite certain that this is the dominating effect. The implications of this and possible mitigation strategies will be discussed in Section \ref{sec:discussion}.

@@

To estimate the B-Modes created by this effect, we have extracted the \emph{Complete Orthogonal Sets of E- and B-Mode Integrals} (COSEBIS, compare \citet{2010A&A...520A.116S}), once of a reference set of correlation functions, to estimate numerical inaccuracies,\footnote{For a reference correlation function the B-Modes should be consistent with zero.} and then for the correlation functions that have been modified to account for a varying depth. We report a consistent B-Mode pattern across all redshift-bins, which can be seen in Figure \ref{fig:bmodes_cosebi}. Although we did not determine the error bars, that a KiDS-like survey would imply on those functions, \citet{2018arXiv181110596A} calculated the COSEBIs of the KV450 survey for the same range in $\theta$. The measured B-Modes were about one order of magnitude higher than the ones created by the varying depth, and still found to be not significant. Given these facts, we conclude that the creation of B-Modes due to varying depth in the KV450 survey is not significant. However, the pattern seems to be very characteristic, so when one encounters B-Modes in next generation surveys, which show a similar pattern, this would suggest that they are created by a similar effect (although we can not exclude other effects that just create the same B-Mode pattern). We also note that the difference in E-Modes is as large as the B-Modes, which suggests that any significant change in the cosmological parameters due to a varying depth will also yield the significant detection of B-Modes.

>>>>

\section{Results}
\label{sec:results}
We applied both our methods to data from the KV450 survey and computed the ratio of observed and modeled correlation functions.\todo{Probably I should include some specifications of the KV450 survey here, like introduce the tomographic bins, etc. Or could I just link to Hendrik's paper?} Furthermore, we conducted numerical simulations investigating the same issue: A $100\,\rm{deg}^2$ field in the Scinet Light Cone Simulations (SLICS) \citep{2018MNRAS.481.1337H} was randomly separated into 10 percentiles. For each tomographic redshift bin of each percentile, galaxies were placed to trace the respective redshift distribution. Afterwards, their expected shear was determined (shape-noise in the form of intrinsic ellipticities of galaxies was not included). This was compared to a set of simulations where the galaxies were simply distributed according to the combined redshift distribution of each respective tomographic bin\todo{Catherine, is this what you did? If you are unhappy with any of these explanations, or feel like I missed something, please feel free to add it :)}. As in \citet{2018arXiv181206076H}, we have separated the data in 5 tomographic redshift bins and performed our analysis for a cross-correlation of all bins. The result can be seen in Figure \ref{fig:all_xis}.
	\begin{figure}
	\centering
	\includegraphics[width=1\textwidth]{images/allxis_0111.pdf}
	\caption{The ratio of observed to modeled correlation functions for the analytic method (green), the semi-analytic method (blue) and the numerical simulations (red) for a cross-correlation of all redshift bins. The numbers in the upper left corners correspond to the respective redshift bins, the upper left triangle depicts the ratios of $\xi_+$, whereas the lower right triangle depicts the ratios of $\xi_-$.}
	\label{fig:all_xis}
	\end{figure}
We can see that for high redshift bins, the analytic and the semi-analytic methods are consistent, whereas for low redshift bins they significantly diverge. We explain this due to the facts that the analyic method uses simplifications that are redshift-dependent and only hold for small variations in redshift, which is not fulfilled in the low redshift bins.\todo{Should I include the plot number density vs average redshift from the talk? Maybe in the appendix to avoid too many Figures?}\footnote{We also observe that in the semi-analytic model, $\xi_-$ seems to be much stronger affected by this effect: Following Equation \eqref{eq:xipm-pkappa}, $\xi_+$ is computed by filtering the power spectrum with the 0-th order Bessel function. This function peaks at $\ell\theta=0$, meaning that for all values of $\theta$, the correlation function $\xi_+$ is sensitive to small values of $\ell$, corresponding to large separations $\theta$. However, $\xi_-$ is obtained by filtering with the 4-th order Bessel function, which peaks at approximately $\ell\theta\approx 5$, so for different $\theta$ this function is sensitive to varying parts of the convergence power spectrum. A more detailed analysis of this can be found in the Appendix of \citet{2017MNRAS.471.4412K}.}

The simulations seem to be in rough agreement with the models, but there are some significant differences. After a thorough analysis we explain these discrepancies the following way: The simulations were performed on a $100\,\rm{deg}^2$ field, which means that shot-noise of the fields plays a significant role. After performing the same simulations for a different distribution of depth between the pointings and obtaining completely different results, we are quite certain that this is the dominating effect. The implications of this and possible mitigation strategies will be discussed in Section \ref{sec:discussion}.

@@

To estimate the B-Modes created by this effect, we have extracted the \emph{Complete Orthogonal Sets of E- and B-Mode Integrals} (COSEBIS, compare \citet{2010A&A...520A.116S}), once of a reference set of correlation functions, to estimate numerical inaccuracies,\footnote{For a reference correlation function the B-Modes should be consistent with zero.} and then for the correlation functions that have been modified to account for a varying depth. We report a consistent B-Mode pattern across all redshift-bins, which can be seen in Figure \ref{fig:bmodes_cosebi}. We also note that the difference in E-Modes is as large as the B-Modes, which suggests that any significant change in the cosmological parameters due to a varying depth will also yield the significant detection of B-Modes.
